Fast, Robust Classification with Applications to Periodic Variable Stars

### 2012-05-10 report/ms.tex
almost done

\documentclass[10pt]{article}
\usepackage{amsmath,textcomp,amssymb,graphicx,amsthm,enumerate}
\usepackage[top=.75in, bottom=.75in, left=.75in, right=.75in]{geometry}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\text{ }}
\newtheorem*{theorem}{Theorem}
\newtheorem{lamma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\begin{document}
\title{Fast, Robust Classification with Applications to Periodic Variable Stars}
\author{Siqi Wu, Mark Rogers, and James Long}
\maketitle

@@

@@

Figure \ref{fig:interval_construction} describes precisely how we subsample the time series and determine interval widths. The details of this procedure could be changed. For example one could bootstrap sample the time measurements, instead of subsampling contiguous sections. Certain computational considerations enter here. It takes around 1 second to derive features for an average length astronomical time series. So sampling, say 20 times, could become prohibitively expansive if the data set is initially at the limits of what is computationally feasible.
\begin{figure*}[ht]
\begin{center}
{\small
\framebox[6in]{
\begin{minipage}[t]{5.8in}
% Insert here your text
\begin{center} \textbf{Constructing Hyper-rectangles from Time Series} \end{center}
\begin{enumerate}
\item Order the light curve measurements in time i.e. $\{(t_1,m_1,e_1),(t_2,m_2,e_2), \ldots , (t_l,m_l,e_l)\}$ where $t_i < t_j$ for $i < j$.
\item Slice each light curve into 5 contiguous sections, producing 5 time series. Slice 1 is $(t_i,m_i,e_i)$ for $0 \leq i \leq .5l$, slice 2 for $.125l \leq i \leq .625l$, slice 3 for $.25l \leq i \leq .75l$ ect.
%% \begin{enumerate}
%% \item $\{(t_1,m_1,e_1), \ldots , (t_{l/2},m_{l/2},e_{l/2})\}$
%% \item $\{(t_1,m_1,e_1),(t_2,m_2,e_2), \ldots , (t_l,m_l,e_l)\}$
%% \end{enumerate}
\item Derive features for each of these slices.
\item Put the interval minimum at the lowest value of the feature for the 5 slices. Put the interval max at the maximum feature value across the 5 slices.
\end{enumerate}
\end{minipage}
}
}
\end{center}
\caption{Description of interval construction algorithm.\label{fig:interval_construction}}
\label{lb}
\end{figure*}

@@

\subsubsection{Performance}
We treat $C$ and $\rho$ as tuning parameters and grid search across them to find the optimal values. We use a grid search for $C$, rather than the path algorithm, because it gives more consistent results. This may be due to a bug somewhere in our path code. In order to assess performance on the grid, we perform 10 fold cross validation, with some initial processing to determine where the grid should be placed.

@@

One attractive feature of using hyper-rectangles to represent uncertainty is the opportunity to obtain sparse solutions. In Figures \ref{fig:RRfeatures} (RR) and \ref{fig:CEPfeatures} (CEP) we list the features and coefficients for all features with non-zero coefficients for the models with the lowest error rate. In both cases, over half the features have coefficients exactly 0. It is also true that in both cases many feature have coefficients which are close to, but not quite 0. We might be able to obtain sparser solutions by increasing our $\rho$. The tradeoff to this, is that this would mean choosing a model that did not obtain optimal misclassification performance.

The performance of the SVM model was compared  all values of $\lambda$ and a grid search on $\rho$.  For a given $(\lambda,\rho)$, the expected loss was computed by performing cross-validation in 90-10 splits so that the union of testing points across splits is the entire dataset and the intersection is disjoint. See surface plots of loss versus $(\lambda,\rho)$: for the CEP dataset, $\lambda\in \lbrace 0\rbrace\cup [0.375,0.750]$ and $\rho\in\lbrace 0\rbrace\cup [0.090,0.032]$ while for the RR dataset, $\lambda\in\lbrace 0\rbrace\cup [54,284]$ and $\rho\in\lbrace 0\rbrace\cup [0.375,0.750]$. Observe that $(\lambda^*,\rho^*)\neq (0,0)$ for both datasets, which implies that the combined model outperforms both of its degenerates. For the CEP dataset, the optimal $(\lambda^*,\rho^*)=(9,0.320)$ yielded 13.25\% classification error while for the RR dataset, $(\lambda^*,\rho^*)=(284,0.48)$ yielded 21.50\% classification error. THESE NUMBERS DON'T LOOK RIGHT, MUST RERUN analyze\_feature.m.

The curves of loss versus $\lambda$ and loss versus $\rho$ were also analyzed.  The loss versus $\lambda$ plot corresponds to the standard point SVM model in Hastie while the loss versus $\rho$ plot corresponds to the interval SVM model in El Ghaoui. Observe the convex-like curvatures, which suggest global minima and hence optimal $\lambda$ and $\rho$.

@@

\subsection{Conclusions}
\begin{enumerate}
\item constructing intervals is non-trivial in many high dimensional problems
\item not clear that this robust framework is ideal when new observation might have low error on certain features
\item kernels, are they possible to use
\item links to code / git repo
\item similarities to elastic net of LASSO / SVM e.g. Wang \cite{wang2007hybrid}
\item our algorithm is not contained (?) in those considered by Russet \cite{rosset2007piecewise}
\end{enumerate}

\subsection{Appendix}
\begin{center}
\begin{figure}[!h]
   \centering
   \includegraphics[width=0.7\textwidth]{./beta0Path.pdf} 
      \caption{The path of the intercept parameter $\beta_0$ for the simulated data. Notice that since $n_+=10>8=n_-$, by the first lemma, $\beta_0(0)=1$.}
   \label{fig:beta0Path}
\end{figure}
\end{center}

\begin{center}
\begin{figure}[!h]
   \centering
   \includegraphics[width=0.7\textwidth]{./alphaPath.pdf} 
      \caption{The path of $\alpha$ for the simulated data.}
   \label{fig:alphaPath}
\end{figure}
\end{center}

>>>>

\documentclass[11pt]{article}
\usepackage{amsmath,textcomp,amssymb,graphicx,amsthm,enumerate}
\usepackage[top=.75in, bottom=.75in, left=.75in, right=.75in]{geometry}
\newcommand{\argmin}[1]{\underset{#1}{\operatorname{argmin}}\text{ }}
\newtheorem*{theorem}{Theorem}
\newtheorem{lamma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\begin{document}
\title{Fast, Robust Classification with Applications to Periodic Variable Stars}
\author{Siqi Wu, Mark Rogers, and James Long}
\maketitle

@@

\begin{figure}[ht]
\begin{center}
\begin{tabular}{cc}
$\begin{array}{cc}
\multicolumn{1}{l}{\mbox{(a)}} & \multicolumn{1}{l}{\mbox{(b)}} \\ \\ \\ [-.35in]
\includegraphics[width=0.45\textwidth]{./beta0Path} &
\includegraphics[width=0.45\textwidth]{./alphaPath}
\end{array}$
\end{tabular}
\end{center}
\caption{(a) Performance across tuning parameters for RR data set. (b) The path of $\alpha$ for the simulated data.}
\label{fig:beta0alphaPath}
\end{figure}

@@

Figure \ref{fig:interval_construction} describes precisely how we subsample the time series and determine interval widths. The details of this procedure could be changed. For example one could bootstrap sample the time measurements, instead of subsampling contiguous sections. Certain computational considerations enter here. It takes around 1 second to derive features for an average length astronomical time series. So sampling, say 20 times, could become prohibitively expensive if the data set is initially at the limits of what is computationally feasible.
\begin{figure*}[ht]
\begin{center}
{\small
\framebox[6in]{
\begin{minipage}[t]{5.8in}
% Insert here your text
\begin{center} \textbf{Constructing Hyper-rectangles from Time Series} \end{center}
\begin{enumerate}
\item Order the light curve measurements in time i.e. $\{(t_1,m_1,e_1),(t_2,m_2,e_2), \ldots , (t_l,m_l,e_l)\}$ where $t_i < t_j$ for $i < j$.
\item Slice each light curve into 5 contiguous sections, producing 5 time series. Slice 1 is $(t_i,m_i,e_i)$ for $0 \leq i \leq .5l$, slice 2 for $.125l \leq i \leq .625l$, slice 3 for $.25l \leq i \leq .75l$ ect.
%% \begin{enumerate}
%% \item $\{(t_1,m_1,e_1), \ldots , (t_{l/2},m_{l/2},e_{l/2})\}$
%% \item $\{(t_1,m_1,e_1),(t_2,m_2,e_2), \ldots , (t_l,m_l,e_l)\}$
%% \end{enumerate}
\item Derive features for each of these slices.
\item Put the interval minimum at the lowest value of the feature for the 5 slices. Put the interval max at the maximum feature value across the 5 slices.
\end{enumerate}
\end{minipage}
}
}
\end{center}
\caption{Description of interval construction algorithm.\label{fig:interval_construction}}
\label{lb}
\end{figure*}

@@

We treat $C$ and $\rho$ as tuning parameters and grid search across them to find the optimal values. We use a grid search for $C$, rather than the path algorithm, because it gives more consistent results. This may be due to a bug somewhere in our path code. In order to assess performance on the grid, we perform 10 fold cross validation, with some initial processing to determine where the grid should be placed.

@@

One attractive feature of using hyper-rectangles to represent uncertainty is the opportunity to obtain sparse solutions. In Figures \ref{fig:RRfeatures} (RR) and \ref{fig:CEPfeatures} (CEP) we list the features and coefficients for all features with non-zero coefficients for the models with the lowest error rate. In both cases, over half the features have coefficients exactly 0. It is also true that in both cases many feature have coefficients which are close to, but not quite 0. We might be able to obtain sparser solutions by increasing our $\rho$. The tradeoff is that this would mean choosing a model that did not obtain optimal misclassification performance.

%% The performance of the SVM model was compared  all values of $\lambda$ and a grid search on $\rho$.  For a given $(\lambda,\rho)$, the expected loss was computed by performing cross-validation in 90-10 splits so that the union of testing points across splits is the entire dataset and the intersection is disjoint. See surface plots of loss versus $(\lambda,\rho)$: for the CEP dataset, $\lambda\in \lbrace 0\rbrace\cup [0.375,0.750]$ and $\rho\in\lbrace 0\rbrace\cup [0.090,0.032]$ while for the RR dataset, $\lambda\in\lbrace 0\rbrace\cup [54,284]$ and $\rho\in\lbrace 0\rbrace\cup [0.375,0.750]$. Observe that $(\lambda^*,\rho^*)\neq (0,0)$ for both datasets, which implies that the combined model outperforms both of its degenerates. For the CEP dataset, the optimal $(\lambda^*,\rho^*)=(9,0.320)$ yielded 13.25\% classification error while for the RR dataset, $(\lambda^*,\rho^*)=(284,0.48)$ yielded 21.50\% classification error. THESE NUMBERS DON'T LOOK RIGHT, MUST RERUN analyze\_feature.m.

%% The curves of loss versus $\lambda$ and loss versus $\rho$ were also analyzed.  The loss versus $\lambda$ plot corresponds to the standard point SVM model in Hastie while the loss versus $\rho$ plot corresponds to the interval SVM model in El Ghaoui. Observe the convex-like curvatures, which suggest global minima and hence optimal $\lambda$ and $\rho$.

@@

\subsection{Conclusions}
We have developed an algorithm and empirically studied an SVM interval model which incorporates feature uncertainty and produces sparse classifiers. The model improves on classification performance over ignoring feature uncertainty. We feel that the results warrent applying this method to larger and multi-class astronomy data sets. From a data perspective we faced several challenges, the greatest of which was determinining intervals for features. This will be common in many high dimensional classification problems where there is uncertainty on features, but no clear way to contruct intervals.
